\input{../../definitions}
\title{\SM{} - Study Guide - Milestone 1 - Task 3: Git, GitHub and the fork-and-branch Git workflow}

\maketitle

\tableofcontents

\lstset{
  showstringspaces=false,
  upquote=true
}

\section{Basic steps}

To work in the
\href{https://github.com/Sistemas-Multimedia/MRVC}{MRMC}
project~\cite{MRVC} you must understand the basics of Git
\cite{Git-book} and \href{https://github.com/}{GitHub}\footnote{There
are other Git-based hosting services such as
\href{https://about.gitlab.com/}{GitLab} and
\href{https://www.atlassian.com/git}{Altassian}/\href{https://bitbucket.org/product}{BitBucket},
but GitHub is the most used one.} \cite{GitHub}, and how to use the
\href{https://guides.github.com/introduction/flow/index.html}{The
  GitHub (Work-)Flow} and the
\href{https://github.com/vicente-gonzalez-ruiz/fork_and_branch_git_workflow}{Fork-and-Branch
  Git Workflow} \cite{fork-and-branch-git-workflow}.

\section{What do I have to do?}

\begin{enumerate}
  
\item Have a look to the \href{https://git-scm.com/book/en/v2}{Pro Git} 
  book \cite{Git-book}, what is the main source of information about
  Git. Keep in mind this reference if you are in trouble with Git.

\item If you don't have an GitHub account, please, do the
  \href{https://guides.github.com/activities/hello-world/}{the Hello
    World guide at GitHub} and create one. Be aware that to contribute
    to MRVC, an GitHub account is required.\footnote{Notice that
    in the Hello World guide,
    the \href{https://git-scm.com/book/en/v2/Git-Branching-Branches-in-a-Nutshell}{\texttt{master}
    branch} is named the \texttt{main} branch.} Please, create
    a \texttt{README.md} file for the Hello
    World \href{https://en.wikipedia.org/wiki/Software_repository}{repo},
    as the guide suggests.

\item Now, we are going to do the same that we would have done using
  the GitHub web interface (except the Step 1: Create a Repo), but now
  using the terminal which will be the most used interface for dealing
  with Git. First, if Git is not installed in your host (try to run
  \texttt{git} in a terminal), install it with:

  \begin{lstlisting}[language=bash]
    sudo apt install git
  \end{lstlisting}

\item (Optional) 
  \href{https://docs.github.com/en/github/getting-started-with-github/github-glossary#clone}{Clone}
  (download) the Hello World repo. You need to click on the ``Code''
  button (select ``https'', not ``download a
  zip file''). Then run:

  \begin{lstlisting}[language=bash]
    cd hello_world
  \end{lstlisting}

  Notice that a new directory named as the repo's name at GitHub has
  been created, and that inside you can find the \texttt{README.md}
  file written in
  \href{https://daringfireball.net/projects/markdown/}{Markdown}.

\item (Optional)
  \href{https://docs.github.com/en/github/getting-started-with-github/github-glossary#checkout}{Create
    (and switch to)} a
  \href{https://docs.github.com/en/github/getting-started-with-github/github-glossary#feature-branch}{feature
    branch} called \texttt{improving\_readme}. In your terminal write:
  
  \begin{lstlisting}[language=bash]
    git checkout -b improving_readme
  \end{lstlisting}
  
\item (Optional) Modify the file \texttt{README.md}. Append to it, for
  example, a link to the Hello World guide. Use an ASCII editor
  (\texttt{nano}, for example):

  \begin{lstlisting}[language=bash]
    nano README
  \end{lstlisting}

  And write:
  
  \begin{lstlisting}
    See the [Hello World](https://guides.github.com/activities/hello-world/) guide at GitHub.
  \end{lstlisting}
  
\item (Optional)
  \href{https://docs.github.com/en/github/getting-started-with-github/github-glossary#commit}{Commit}
  your modification(s):

  \begin{lstlisting}[language=bash, showstringspaces=false, upquote=true]
    git commit -am "Providing the Hello World link"
  \end{lstlisting}

  In your first commit you will be prompted with:

  \begin{lstlisting}[language=bash]
    git config --global user.email "your_email@example.com"
  \end{lstlisting}

  Please, input such information.
  
  After the \texttt{commit}, your \href{https://docs.github.com/en/github/getting-started-with-github/github-glossary#fetch}{\emph{local}} repo is \emph{ahead} of
  your
  \href{https://docs.github.com/en/github/getting-started-with-github/github-glossary#origin}{\emph{origin}}
  (copy at GitHub of the) remote repo. This means that your
  \emph{local} has modifications that the \emph{origin} doesn't have yet.

\item (Optional) Synchronize your \emph{local} and the \emph{origin}
  using
  \href{https://docs.github.com/en/github/getting-started-with-github/github-glossary#push}{push}:

  \begin{lstlisting}[language=bash]
    git push
  \end{lstlisting}

  Notice that if you have not uploaded a public
  \href{https://www.ssh.com/}{SSH}
  \href{https://www.ssh.com/ssh/identity-key}{key} (or the
  corresponding private key is not properly installed in your
  computer), the GitHub server requests your username and password,
  and this is something that is going to happen with every
  \texttt{push}. To avoid this repetitive input of your GitHub login
  information, you need
  \href{https://docs.github.com/en/github/authenticating-to-github/connecting-to-github-with-ssh}{to
    login at GitHub} using
  \href{https://en.wikipedia.org/wiki/Public-key_cryptography}{public-key
    criptography}. For that, you must own a pair of keys, one public
  and other private, and upload the public one to GitHub.
  
\item The first step is to check whether you already have a pair of
  keys (if your are using the just installed Xubuntu distribution,
  obviously you don't need to check anything and can go directly to
  the next step). Simply revise your \texttt{\$HOME/.ssh} directory
  with:

  \begin{lstlisting}[language=bash]
    ls -l ~/.ssh
  \end{lstlisting}

  and if you find a pair of files with almost the same name, and one
  of they finising in \texttt{.pub}, you probably own a pair of SSH
  keys.

\item If you don't have one or if you prefeer to create a new one,
  you can create a pair of keys. Open a terminal and write:

  \begin{lstlisting}[language=bash]
    ssh-keygen -t rsa -b 4096 -C "your_email@example.com"
  \end{lstlisting}

  using the email address you provided when you created your GitHub
  account. Then, when you are prompted with:

  \begin{lstlisting}[language=bash]
    Enter a file in which to save the key (/home/you/.ssh/id_rsa):
  \end{lstlisting}

  just press the Enter-key, to select such output prefix. Otherwise,
  write a different one, but don't change the path to the
  \texttt{.ssh} directory.

\item Now SSH should request you for a
  \href{https://en.wikipedia.org/wiki/Passphrase}{passphrase}. If you
  write one, you will be asked for it each time you push your commits
  to GitHub. There are two options to avoid this:

  \begin{enumerate}
  \item Input no passphrase (just by pressing the Enter-key again in
    the previous step). This has the drawback that if somebody steals
    your keys, he could access to GitHub as he were you.
  \item Input a passphrase and configure
    \href{https://www.ssh.com/ssh/agent}{\texttt{ssh-agent}} to send
    it to GitHub by you. This option is the preferable one because you
    will be asked for the passphrase only when the \texttt{ssh-agent}
    is started (\href{https://www.xfce.org/}{Xfce} does that for you).
  \end{enumerate}

\item Now it's time to check whether the \texttt{ssh-agent} is already
  running in your computer. This can be done with:

  \begin{lstlisting}[language=bash]
    ps aux | grep ssh-agent
  \end{lstlisting}

  and in the case of Xubuntu, you should get something similar to:

\begin{verbatim}
 989 ?        Ss     0:00 /usr/bin/ssh-agent /usr/bin/im-launch startxfce4
1433 pts/0    S+     0:00 grep --color=auto ssh-agent
\end{verbatim}

  This means that there are two
  \href{https://en.wikipedia.org/wiki/Process_(computing)}{processes}
  in whose description there exists the string \texttt{ssh-agent}. The
  first entry is the agent process. The second one is the
  \texttt{grep} running at the same time that the
  \href{https://man7.org/linux/man-pages/man1/ps.1.html}{\texttt{ps}}.

\item If the \texttt{ssh-agent} were not running, it can be launched
  to run in
  the \href{https://en.wikipedia.org/wiki/Background_process}{background}
  with:

  \begin{verbatim}
    eval "$(ssh-agent -s)"
  \end{verbatim}

  but you don't need to do that in your Xubuntu installation, because
  (remember) the \texttt{ssh-agent} the
  Xfce desktop environment lauches it.

\item With your keys, run:

  \begin{lstlisting}[language=bash]  
    ssh-add ~/.ssh/id_rsa
  \end{lstlisting}

  and the passphrase will be prompted. Notice that you specify the
  private key, not the public one.
  
\item Go now to GitHub \texttt{->} Settings \texttt{->} SSH and GPG
  keys \texttt{->} New SSH key. Open a terminal and write:
  
  \begin{lstlisting}[language=bash]  
    cat .ssh/id_rsa.pub
  \end{lstlisting}

  and copy and paste the content of such file (which ends with your
  email address) inside of the space where you can read
  ``\texttt{Begins with 'ssh-rsa', ...}''. Don't forget to give a
  title (something such as ``\texttt{sm}'' (sistemas multimedia))
  to the key pair if you created a dedicated one.
  
\item When you use the key for the first time (clonning a repo or
  pushing a commit), the SSH client will warn you that the autenticity
  of \texttt{github.com} cannot be checked. This is normal and should
  happen only in the first interaction. Type \texttt{yes}. Notice that
  if this problem persists, then you could be suffering a
  \href{https://en.wikipedia.org/wiki/Man-in-the-middle_attack}{man-in-the-middle
    attack}.

\item Revise
  \href{https://github.com/vicente-gonzalez-ruiz/fork_and_branch_git_workflow}{The
    Fork and Branch Git Workflow}. Basically, this ``protocol''
  explains that to contribute to an open-source repo hosted by GitHub
  without belonging to the developping
  \href{https://docs.github.com/en/organizations/collaborating-with-groups-in-organizations/about-organizations}{organization},
  you must do some \texttt{git}-steps (and some of them are below).

\item Make a fork of the
  \href{https://github.com/Sistemas-Multimedia/MRVC}{MRVC}
  project. We will call to this repo the
  \href{https://docs.github.com/en/github/getting-started-with-github/github-glossary#upstream}{\emph{upstream}}, whose URL is
\begin{verbatim}
  git@github.com:Sistemas-Multimedia/MRVC.git
\end{verbatim}        
  This info can be found when you
  \href{https://docs.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository}{clone}
  MRVC. Notice, that the action of clonning (not forking) the MRVC is a waste of
  time because you cannot contribute directly to it (remember, you
  must \texttt{clone} your own repo).

\item Add the
  \href{https://docs.github.com/en/github/getting-started-with-github/github-glossary#remote}{remote}\footnote{Git
  is a
  \href{https://nvie.com/posts/a-successful-git-branching-model/}{decentralized
    control system for source code}. Decentralization means that every
  developer has a copy of the \emph{origin}, and that, the developers
  can synchronize their \emph{locals} with any of the \emph{remotes}.}
  \emph{upstream} with:

  \begin{lstlisting}[language=bash]
    git remote add upstream git@github.com:Sistemas-Multimedia/MRVC.git
  \end{lstlisting}

  Check that everything has worked with:

  \begin{lstlisting}[language=bash]  
    git remote -v
  \end{lstlisting}

  where you should see two remotes: \emph{origin} and \emph{upstream}. Something similar to:

\begin{verbatim}
origin   git@github.com:you_at_GitHub/MRVC.git (fetch)
origin   git@github.com:you_at_GitHub/MRVC.git (push)
upstream git@github.com:Sistemas-Multimedia/MRVC.git (fetch)
upstream git@github.com:Sistemas-Multimedia/MRVC.git (push)
\end{verbatim}

\item To update your local repo write:

\begin{verbatim}
git pull upstream master
\end{verbatim}

\end{enumerate}

\section{Some extra (and optional) configuration}

\begin{verbatim}
cat >> ~/.profile << EOF

# Add branch name to the Bash prompt:
parse_git_branch() {
    git branch 2> /dev/null | sed -e '/^[^*]/d' -e 's/* \(.*\)/ (\1)/'
}
export PS1="\[\033[01;32m\][\u@\h\[\033[01;37m\] \W\[\033[01;32m\]]\$(parse_git_branch)\[\033[00m\]\$ "

EOF
\end{verbatim}

\section{Timming}

Please, finish this task before the next class.

\section{Deliverables}

None.

\section{Resources}

\renewcommand{\addcontentsline}[3]{}% Remove functionality of \addcontentsline
\bibliography{git,MRVC}
