% Emacs, this is -*-latex-*-

% https://vcgit.hhi.fraunhofer.de/jvet
% https://www.etsist.upm.es/uploaded/392/02_Video_Compression_Basics_MPEG1_MPEG2_MPEG4.pdf

\input{../../definitions}
\title{Temporal Transforms}

\maketitle
\tableofcontents

\section{Temporal correlation}
%{{{

In general, neighbor \emph{frames} (or images) in (video)
\emph{sequences} exhibit a high temporal correlation degree that can
be exploited to improve significantly the RD curves. This correlation
produces a temporal redundancy that can be removed using (temporal)
decorrelating techniques. Such techniques can be considered as a
special type of transform applied to the temporal domain: temporal
transfors.

A temporal transform inputs two or more frames\footnote{With pixels or
  coefficients, depending on the current domain in which the frame has
  been represented.}, and outputs at least one \emph{residual} (frame)
in which the residual pixels have a higher dynamic range but, in
general, also a high energy concentration.

%}}}

\section{Motion Compensation (MC)}
%{{{

Most video coding standards use
\href{https://en.wikipedia.org/wiki/Motion_compensation}{Motion
  Compensation (MC)} to generate the residual
frames~\cite{vruiz__MC}. MC exploits the temporal correlation and
reduces the entropy of the residuals\footnote{The better the
prediction, the lower the entropy of the residuals.}. Basically, MC
consists in subtracting from each original frame a prediction (frame)
built with the information that must be also avaliable\footnote{In
order to make a reversible process.} at the decoder. Notice that,
after using MC, the number of pixel in the residual frame is equal to
the number of pixels in the compensated frame.\footnote{At least, when
we compensate in the image domain.}

%}}}

\section{Motion Estimation (ME)}
%{{{

To compensate the motion we need first to estimate\footnote{In most of
the situations, the determination of the true motion of the objects in
a real scene is a ill-posed problem because it is impossible to find
it using only a sequence of 2D images. A different situation is when
we use al least 2 cameras.} it using Motion Estimation (ME)
techniques~\cite{vruiz__ME}. Using the motion fields generated by the
motion estimator (and obviously, the pixel data), both the encoder
and the decoder generate the predictions that will be added (by the
decoder) from the predicted images~\cite{vruiz__MC}. However, notice
that in most of the video coding standards, ME is only performed by
the encoder because it is a costly operation, and for this reason, the
motion vector fields must be transmitted to the decoder. This responds
to the idea of ``compress one, decompress many''.

%}}}

\section{GOF-ing}

The RD performance of ME/MC depends on the amount of temporal
redundancy in the sequence. If such an amount is low, it can be more
RD-efficient to interrupt the (ME/)MC process. The set of consecutive
frames in which MC is active is usually known as a GOF\footnote{Some
standards also use GOP (Group Of Pictures).} (Group of Frames). Notice
that, under the RD prism, the GOF partitioning (the length of the
GOFs) should be an adaptive process controlled by a RD control
algorithm.

However, in some contexts\footnote{Specifically, constant bit-rate
  encodings.} it may be necessary to use a fixed GOP
partition~\cite{vruiz__MC}. For example, if we want to give the option
to the users to move fast forward or backward along the sequence, we
need to set some given GOF size. Another reason to use a constant GOF
size is to limit the propagation of decoding errors (for example,
because in a streaming session we have not received some data). When a
new GOF starts, the propagation of such errors vanishes.

\section{Block-based MC and Rate/Distortion Optimization (RDO)}

The MC schemes used in most video coding standards compensate blocks
of pixels~\cite{vruiz__ME}. In this context, depending on the block
decision mode implemented in the RDO procedure\footnote{The part of
the RDO procedure that controls the block-type.}, blocks can be of
different type (I (intra), P (predicted), B (bidirectionally
predicted) and S (skipped))~\cite{vruiz__MC}. A I-block is used when,
for this block, we do not found enough temporal correlation between
frames and from a RD perspective, it is more advantagous to use
\emph{intra-coding}. When we found one (P) or more (B) reference
blocks to perform a good prediction, we are using
\emph{predictive-coding}. Notice that the number of reference blocks
can be higher than one, a number that also should be controlled by
RDO. In the intra-coding mode, all the frames are I-type because
otherwise we could not reset the propagation errors. In Motion
Compensated Temporal Filtering~\cite{vruiz__MCTF}, the frames can also
be I, P or B.

\section{Frame types}

Depending on the type of blocks used in the frames, we have different
types of frames: I, P, and B~\cite{vruiz__MC}. In a I-frame, all
blocks are I-type. In a P frame, I- and P-type blocks can be found. In
a B-frame all types of blocks can be used.

\section{Resources}
\begin{enumerate}
\item \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/IPP_DCT.ipynb}{IPP\_DCT.ipynb} in VCF.
\item \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/MCTF.ipynb}{MCTF.ipynb} in VCF.
\item \href{https://github.com/vicente-gonzalez-ruiz/motion_estimation/blob/main/src/motion_estimation/full_search_block_ME.ipynb}{Full search block-based ME (Motion Estimation)}.
\item \href{https://github.com/vicente-gonzalez-ruiz/motion_estimation/blob/main/src/motion_estimation/full_search_dense_ME.ipynb}{Full search dense (1x1) ME}.
\item \href{https://github.com/vicente-gonzalez-ruiz/motion_estimation/blob/main/src/motion_estimation/farneback_ME.ipynb}{Farnebäck's motion estimation}.
\item \href{https://github.com/vicente-gonzalez-ruiz/motion_compensation/blob/master/introducing_IPPP.ipynb}{Introducing the Low-delay (IPP...) Mode}.
\item \href{https://github.com/Sistemas-Multimedia/MRVC}{Multi-Resolution Video Coding (MRVC)}.
\end{enumerate}

\section{To-Do}
%{{{

\begin{enumerate}
  
\item The module used in
  \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/IPP_DCT.ipynb}{IPP\_DCT.ipynb}
  should determine the block type (I, P, or B) using the number of
  bits that actually the block requires after cuantization (now an
  heuristic using the variance of the quantized block is
  implemented). Notice that this implies that the blocks must be
  entropy encoded independently (the codec must operate at the block
  level, not at the image one). For this, arithmetic coding is highly
  recommended\footnote{Notice that we can estimate the number of bits
  that a block can require when we encode it with a a 0-order
  context-based arithmetic using the entropy of the data of the
  codec. If we have several context, the entropy can be computed
  context-based also.}.

\item Both,
  \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/IPP_DCT.ipynb}{IPP\_DCT.ipynb}
  and
  \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/MCTF.ipynb}{MCTF.ipynb}
  estimate the motion at the frames resolution. Better predictions can
  be generated using sub-pixel motion estimation \cite{timoner2000subpixel,farneback02}.

\item Both,
  \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/IPP_DCT.ipynb}{IPP\_DCT.ipynb}
  and
  \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/MCTF.ipynb}{MCTF.ipynb}
  load first the complete video in memory to process it. This is a
  problem when the videos are long or have a high resolution. Work at
  the frame level. Disks are faster enought.

\item In the current implementation of
  \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/MCTF.ipynb}{MCTF.ipynb}
  the reference frames have the maximum quality when encoding, and
  this is not the case with the video is decoded (obviously, if the
  QSS is bigger than 1). To avoid the drift error along temporal
  resolution levels, use at the encoder the same references that the
  decoder will use. Notice that this implies that the quantization
  step size must known at compression time.
  
\item Create a video codec similar to the used in
  \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/IPP_DCT.ipynb}{IPP\_DCT.ipynb},
  but using an ``IBP...'' scheme. In this case, the predicted images B
  use more than one reference image \cite{vruiz__ME,vruiz__MC}. For
  now, use only the adjacent frames as reference.
  
\end{enumerate}

%}}}

\section{References}
%{{{

\renewcommand{\addcontentsline}[3]{}% Remove functionality of \addcontentsline
\bibliography{image_pyramids,DWT,motion_estimation,video_compression}

%}}}
