% Emacs, this is -*-latex-*-

% TODO: In the DWT domain, use block-based neural compression using autoencoders

% https://stackabuse.com/autoencoders-for-image-reconstruction-in-python-and-keras/
% https://github.com/FireFYF/modulatedautoencoder
% https://www.tensorflow.org/tutorials/generative/data_compression
% https://www.youtube.com/watch?v=x_q7cZviXkY ( PCS 2018 – Learned Image Compression )
% https://keras.io/examples/generative/vq_vae/
% https://jhui.github.io/2017/03/06/Variational-autoencoders/
% https://towardsdatascience.com/difference-between-autoencoder-ae-and-variational-autoencoder-vae-ed7be1c038f2
% https://deepai.org/machine-learning-glossary-and-terms/prior-probability
% https://towardsdatascience.com/auto-regressive-generative-models-pixelrnn-pixelcnn-32d192911173
% https://openaccess.thecvf.com/content_cvpr_2018_workshops/papers/w50/Le_Tan_DeepVQ_A_Deep_CVPR_2018_paper.pdf
% https://www.deeplearningbook.org/
% https://arxiv.org/pdf/1711.00937.pdf (Neural Discrete Representation Learning)
% https://colab.research.google.com/github/keras-team/keras-io/blob/master/examples/generative/ipynb/vq_vae.ipynb (Vector-Quantized Variational Autoencoders)
% https://www.tensorflow.org/tutorials/generative/autoencoder
% https://blog.paperspace.com/autoencoder-image-compression-keras/
% https://www.analyticsvidhya.com/blog/2021/06/complete-guide-on-how-to-use-autoencoders-in-python/
% https://colab.research.google.com/github/keras-team/keras-io/blob/master/examples/generative/ipynb/vae.ipynb
% https://blog.keras.io/building-autoencoders-in-keras.html
% https://keras.io/examples/generative/vae/
% https://arxiv.org/pdf/1611.01704.pdf (END-TO-END OPTIMIZED IMAGE COMPRESSION)
% https://www.tensorflow.org/tutorials/generative/data_compression
% https://www.researchgate.net/profile/Asma-Salem-4/publication/319764377_Image_Compression_Using_Wavelet_and_Subband_Based_Transform/links/59bbe3d2458515e9cfc79d0d/Image-Compression-Using-Wavelet-and-Subband-Based-Transform.pdf
% https://ocw.mit.edu/courses/18-06-linear-algebra-spring-2010/resources/lecture-31-change-of-basis-image-compression/
% https://arxiv.org/abs/1908.08988 (Neural Image Compression and Explanation)
% https://arxiv.org/abs/2011.03029 (CompressAI: a PyTorch library and evaluation platform for end-to-end compression research)
% https://heartbeat.comet.ml/a-2019-guide-to-deep-learning-based-image-compression-2f5253b4d811
% https://scelesticsiva.github.io/assets/image_compression_report.pdf
% https://interdigitalinc.github.io/CompressAI/intro.html
% https://arxiv.org/abs/1802.01436 (Variational image compression with a scale hyperprior)
% 

\input{../../definitions}
\title{\href{https://sistemas-multimedia.github.io/contents/spatial_transforms/}{Spatial Transforms}}

\maketitle

\tableofcontents

\section{Spatial decorrelation}
%{{{

Spatial transforms used in image and video compression exploit the
statistical correlation (and perceptual redundancy\footnote{We will
  see this with more detail later in this course.}) that the pixels
show as a consequence of the spatial (2D) correlation that is present
in most of the images (and video frames). For example, some textures of
an image can occur more than once (in the same image). Also, usually
happens that neighbor pixels have similar values.

While color transforms are pixel-wise computations, spatial transforms
are image-wise (or at least, block-wise). This means that the
transform inputs an image (of pixels) and outputs a matrix of
coefficients, which generally express the resemblance between the
image and a set of basis functions, usually orthogonal. For example,
after using the 2D-DCT (two-dimensional Discrete Cosine
Transform)~\cite{vruiz__DCT}, (the index of) each coefficient
represents a different spatial frequency\footnote{That depends on the
  position of the coefficient in the transformed domain.} and its
value, the amount of the corresponding basis found in the image. In
the case of the dyadic 2D-DWT~\cite{vruiz__DWT}, the coefficients
``speak'' additionally about a spatial resolution in the image pyramid
and position inside of the pyramid level.

%}}}

\section{Benefits of spatial transforms}
%{{{

Most spatial transforms provide:
\begin{enumerate}
\item \textbf{Energy concentration:} Usually, a small set of
  (low-frequency) coefficients represents most of the information
  (energy) of the image. This decreases the entropy\footnote{When the
    entropy is decreased while the information is preserved, this
    usually means that an entropy encoding algorithm will perform
    better.} and increases the range of the quantization step sizes,
  because the dynamic range of the coefficients is higher than the
  dynamic range of the pixels\footnote{Quantization is a discrete
    operation constrained by the number of bits used to represent the
    quantization indexes. When the dynamic range of a signal is high,
    this makes possible to use more quantization levels and therefore,
    a higher number of available RD points.}.
\item \textbf{Low/High frequency analysis:} The human visual system is
  more sensitive to the low frequencies (for this reason, the
  \href{https://en.wikipedia.org/wiki/Contrast_(vision)#Contrast_sensitivity}{contrast
    sensitiviy function} is not flat). This means that we can quantize
  more severely the high frequencies without generating a perceptible
  distortion.
\item \textbf{Multiresolution:} Depending on the transform, it is
  possible to reconstruct the original image by resolution
  levels~\cite{vruiz__DWT}. This option can be interesting when the
  resolution at which the image must be reconstructed is not known a
  priori. For example, JPEG 2000 (which is based on the 2D-DWT) is
  used in \href{https://jpeg.org/jpeg2000/applications.html}{digital
    cinema} because, although movie players do not have the same
  resolution in all movie theaters, the same code-stream (with the
  maximum resolution) can be used in all of them.
\end{enumerate}

%}}}

\section{Scalar quantization and rate-distortion optimization in the transform domain}
%{{{

Scalar quantization is efficient in the transform domain because the
coefficients are decorrelated. The next logical step (after
quantization) is the entropy coding of the quantization indexes. Here,
depending on how the coefficients are quantized, we can trace
different RD curves (all of them starting (and finising) in the same
(rate, distortion) point). For example, if we compress each
subband\footnote{In the case of a spatial transform, a subband is form
  by all the coefficients that describe the same frequency components
  in different areas or resolutions (when available) of the image.}
independently, we must find the quantization step sizes that select
the same slope in the RD curve of each
subband~\cite{vruiz__information_theory}. A RD curve is a discrete
convex function where each line that connects adjacent RD points has a
slope.  In this context, given a target distortion or rate, the
quantization step size used in each subband should generate the same
slope.

%}}}

\section{Block-based transforms}

Depending on the content of an image and the spatial transform, it can
be necessary to divide the image into 2D chunks (usually called
\emph{tiles} or \emph{blocks}), and encode each chunk
independently. In general:
\begin{enumerate}
\item Tiles are used when the image is made up of very different areas
  (for example, with text and natural picture content). Tiles are
  usually rectangular but can have any size, and are usually defined
  attending to RD or perceptual issues (for example, text is not well
  compressed by lossy codecs).
\item Blocks are smaller than tiles and, in most of cases,
  squared. The block partition can be adaptive and, in this case,
  should be found using RDO (Rate Distortion Optimization)\footnote{If
    no other more important requirement exists, such as
    multiresolution.}.
\end{enumerate}

\section{``Hand-crafted'' VS learned transforms}
%{{{

Using machine learning techniques (for example, with an artificial
neural network), it is possible to build a machine-krafted transform,
specifically tuned for some type of images, or at least, capable of
determining specific features in the images. This technique receive
the name of analysis–synthesis dictionary learning, linear
autoencoders, or learned transform coding.

Potentially, learned-based image (and video) compressors are adaptive
algorithms that can be more efficient than those in which the
transforms are pre-defined.

%}}}

\section{Linearity}
Linear systems satisfy the Superposition Principle that basically sais
that the transform of linear combination of signals is equal to the
linear combination of the transform of the signals. This makes the
systems predictable and easier to analyze. Thus for example, if we
double the input of a linear system we now that the output will be
also doubled.

\section{Orthogonality}
In Transform Coding, (among other reasong) orthogonality is
interesting because the resulting coefficients will not share
information (in terms the similarity between the input signal and the
basis functions). This maximizes energy accumulation in the transform
domain. Orthogonality implies linearity.

\section{Filter symmetry}
From a signal processing point of view, the basis functions that
conform the transform matrices can also be considered as filters. When
the filters are symmetric (linear in phase) all the filters of the
transform generate the same shift of the phase of the signal, but when
the filter are asymmetric this does not happen and ringing artifacts
are generated at the edges of the objects of the encoded image.

\section{Bi-orthogonality}
It is mathematically proven that you cannot have a wavelet that is
finite, orthogonal, and symmetric all at the same time (with the
exception of the very blocky Haar wavelet). Bi-orthogonal transforms,
on the other hand, can be created using finite-support symmetric
filters at the cost of using filters with different gains that usually
are nor orthogonal. A bi-orthogonal transform can be easely recognized
because the synthesis matrix is not the transpose of the analysis
matrix.

\section{Orthonormality}
An orthonormal transform is a orthogonal transform that satisfy that
after using the analysis and the synthesis transform, we obtain
exactly the same input signal. In a orthogonal transform, the output
could be amplified (depending on the transform)

\section{Separability}
A separable transform is those in which the transform of a
multidimensional signal is equal to the unidimensional transform
applied to each dimension.

\section{Gaussian sources and the Central Limit Theorem}
In Gaussian signals, the samples follow a Gaussian distribution. In
the real world, signals are rarely perfectly perfectly
mathematical. However, the Central Limit Theorem states that when you
add many independent random signals together, their sum naturally
gravitates toward a Gaussian distribution.

\section{Discrete Cosine Transform (DCT)}
The DCT is orthonormal and separable. It is usually applied by blocks
of 8x8 pixels because:
\begin{enumerate}
\item In most natural images the spatial correlation is only
  significative in small regions (we are not going to gain in
  compression ratio using the full-resolution DCT).
\item It is faster: the DCT requires $N\log_2(N)$ operations while
  blocking only $N$.
\end{enumerate}

In VCF has been implemented the
\href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/2D-DCT.ipynb}{2D-DCT}.

\section{Modified DCT (MDCT)}
To avoid blockiness we can overlap the blocks. The MDCT was
speficically designed to avoid this artifact. See
\href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/2D-MDCT.ipynb}{2D-MDCT}.

\section{Karhunen–Loève Transform (KLT)}
In the KLT the basis funtions are the eigenvectors of the covariance
matrix of the input data. In terms of energy concentration (also
called
\href{https://web.stanford.edu/class/ee398a/handouts/lectures/07-TransformCoding.pdf}{transform
  coding gain} \cite{sayood2017introduction}), for a zero-mean
Gaussian sources the KLT is optimal among all orthogonal transforms.

Like the other transforms studied here, the KLT is orthonormal but in
general not separable, and applied block-by-block for two main
reasons:
\begin{enumerate}
\item The basis (eigenvectors) are signal dependent and therefore, the
  decoder needs the inverse transform matrix, whose size is $N^2$ for
  an input of $N$ pixels ($N$ eigenvectors of length $N$). Remember
  that the KLT is not separable.
\item Most natural images are non-stationary, meaning their statistics
  change drastically from one area to another (the top half of a photo
  might be a smooth, highly correlated blue sky, while the bottom half
  might be highly detailed, uncorrelated grass). Therefore, if you
  apply a global KLT to the whole image, it creates a set of basis
  functions that average out the sky and the grass. This "average"
  basis isn't optimal for either region.
\end{enumerate}

A 2D-blocks version of the KLT can be found in
\href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/2D-KLT.ipynb}{2D-KLT}. Notice
that only a single analysis matrix has been used when encoding the
whole image (and only a synthesis matrix can be used for decoding).

\section{Discrete Wavelet Transform (DWT)}
The DWT is linear, separable and at least bi-orthogonal. Compared to
the DCT, in the DWT the basis functions have a local support (are not
infinite as the cosine functions), and the DWT coefficients express
both frequency and locality (in the original image). This has two main
advantages:
\begin{enumerate}
\item The DWT can be applied without using blocks, avoiding the
  blocking artifact usually visible at very low bit-rates.
\item In general, the energy compactation generated by the DWT is
  superior to the achieved by the DCT.
\end{enumerate}

VCF implements
\href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/2D-KLT.ipynb}{2D-DWT}.

\section{RD-optimized Learned Block Transform (LBT)}
Neural networks are good learning things. Specifically, autoencoders
are capable of matching correspondences between input and output
tensors. If we design a 2-layers autoencoder where the neurons do not
use any activation function and we train it to minimize the cost
function
\begin{equation}
  J = D + \lambda R,
\end{equation}
where $D$ represents distortion and $R$ bit-rate, we will build a
Rate/Distortion (RD) optimized transformation. Notice that, unlike the
KLT, now we are considering the quantization and the entropy coding
subsystems (and their parameters, such as the quantization step size
and the order of the context modeling) to minimize the $J$ tradeoff
between $D$ and $R$ (in the KLT we only minimize $D$). $\lambda$
controls our preference:
\begin{enumerate}
  \item If $\lambda$ is small, for a given quantization step size and
    entropy encoding parameters, we are interested in finding a
    transformation matrix that should be efficient for high quality
    reconstruction, at the expense of increasing the bit-rate.
  \item If $\lambda$ is big, you are telling to the optimizer:
    ``Bit-rate is expensive''. The autoencoder will update its weights
    to learn a transform that aggressively groups and compress
    features (``transform coefficients'', in this case) even a the
    cost of a high distortion. Therefore, the transform will be
    optimized for low bit-rates.
\end{enumerate}

In VCF has been implemented a block-wise
\href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/2D-LBT.ipynb}{2D-LBT}.

\section{To-Do}
\begin{enumerate}
  \item The current implementation of LBT is not separable (the
    transform matrices has been found for the complete blocks). If
    $T(\mathbf{X})$ represents the transform of the block $\mathbf{X}$
    modify
    \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/notebooks/2D-LBT.py}{2D-LBT}
    to train $\mathbf{A}\in\mathbb{R}^{B\times B}$, where $B$ is the
    side of the square blocks, being\footnote{In the current
    implementation,
    \begin{equation}
      T(\mathbf{X}) = \mathbf{B}\mathbf{X},
    \end{equation}
    where $\mathbf{B}\in\mathbb{R}^{B^2\times B^2}$.}
    \begin{equation}
      T(\mathbf{X}) = \mathbf{A}\mathbf{X}\mathbf{A}^T.
    \end{equation}
    
    Sumarizing, the transform is applied first by rows and the by
    columns. The transform matrix is smaller.

  \item In the current implementation
    \begin{equation}
      R = \frac{1}{N}\sum_{i=1}^{N} \log(\sigma_i^2),
    \end{equation}
    where $N$ is the number of pixels/block and $\sigma^2_i$ is the
    variance of the $i$-the coefficient inside of each block.
    Obviously, $R$ is only an estimation of the (inverse of) real
    bit-rate, which depends also on the QSS and the entropy encoder
    parameter(s). Try to improve the accuracy of this term of the cost
    function. Modify any other training parameter (such as the number
    of epochs) to achive this goal.
    
\end{enumerate}

\section{Resources}
%{{{

\begin{enumerate}
\item
  \href{https://github.com/vicente-gonzalez-ruiz/DCT2D/blob/master/src/DCT2D/YCoCg_2D_DCT_SQ.ipynb}{Image
    Compression with YCoCg + 2D-DCT}.
\item
  \href{https://www.tensorflow.org/tutorials/generative/data_compression}{Learned
    data compression}.
\item
  \href{https://github.com/vicente-gonzalez-ruiz/learned_image_compression/blob/main/notebooks/LIC.ipynb}{Learned
    Image Compression (LIC) using auto-encoders}.
\item
  \href{https://github.com/vicente-gonzalez-ruiz/learned_image_compression/blob/main/notebooks/AutoencoderBlockCompression.ipynb}{AutoencoderBlockCompression.ipynb}.
\item
  \href{https://github.com/fchollet/deep-learning-with-python-notebooks}{Companion
    Jupyter notebooks for the book ``Deep Learning with
    Python''}~\cite{chollet2021deep}.
\end{enumerate}

%}}}

\begin{comment}
\section*{To-Do}
%{{{

\begin{enumerate}
\item Add a new option to
  \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/src/2D-DWT.py}{the
    VCF codec \texttt{2D-DWT.py}}, that throught RDO, determine the
  optimal DWT basis (defined in
  \href{https://pywavelets.readthedocs.io/en/latest/}{PyWavelets}). \textbf{2
    points}.
\item Create a new VCF image codec (similar to
  \href{https://github.com/Sistemas-Multimedia/VCF/blob/main/src/2D-DCT.py}{\texttt{2D-DCT.py}})
  where the 2D-DCT is replaced by an ANN (artificial neural
  network). The forward transform should maximize the concentration of
  energy, and the inverse transform should restore the original
  block. To avoid transmit the coefficients of the ANN, use a model
  previously trained to compute the 2D-DCT, and retrain the model
  using only the previously encoded block of the image. For example, when
  encoding the second block, use the first block to fine-tune the
  model. For the third block, use the second block to retrain the
  model, and so on. \textbf{10 points}.
\item Create a new image codec (similar to \texttt{2D-DWT.py}) where
  the coefficients of the filters used in the 2D-DWT are determined by
  an ANN trained to maximize the transform gain (energy
  concentration). Initially use the coefficients of the filters of a
  well-known DWT kernel, and in subsequent iterations (levels) of the
  transform used the previously analyzed spatial resolution levels to
  fine-tune the coefficients. \textbf{10 points}.
%\item If the right coefficients in the DWT domain are multiplied by a
%  power of 2 before quantization, we can define a ROI of arbitrary
%  shape at compression time~\cite{vruiz__JPEG2000}. Include this
%  possibility in the DWT-based compression pipeline of VCF.
% \item Modify CVF to use SPIHT.
  % \item LPT can replace the DWT. Can we travel from LPT to DWT using any filter?
  % \item Irregular Subsampling + normalized convolution (ver tesis Gunnan Farneback).
\end{enumerate}
%}}}
\end{comment}

\section{References}
%{{{

\renewcommand{\addcontentsline}[3]{}% Remove functionality of \addcontentsline
\bibliography{projects,signal_processing,neural_nets,information_theory,JPEG2000,data_compression}

%}}}

\begin{comment}

  
\section{Tight Laplacian Pyramid Transform (TLPT)}
%{{{

Pyramid frames were introduced in 1983 by Burt and Adelson. In some
way, they can be considered one of the precursors of the wavelet
octave band decompositions.

Technically, the LPT is a frame expansion that generates an expanded
octave-band decomposition. The ratio of redundancy in the
decomposition $R\rightarrow 2$ with the number of levels.

Because of the shift invariance, MC can be incorporated in the enhancememt layer.

The LP with orthogonal filter is a tight frame.

%}}}

\section{Critially Sampled LPT (CS-LPT)}
%{{{

Technically, the LPT is a frame expansion, that generates an expanded
octave-band decomposition. The ratio of redundancy in the
decomposition $R\rightarrow 2$ to the number of DWT levels.

The 2D-DWT can be used on sequences of images, by simply iterating as described in the Algorithm~\ref{alg:MDWT}. $I$ is the number of
images in the sequence $V$.

\begin{pseudocode}{\text{MDWT}}{V}
  \label{alg:MDWT}
  \FOR i \GETS 0 \TO I-1 \DO
  V_i = \text{2D-DWT}(V_i)
\end{pseudocode}

%}}}

\section{Autoencoder Latent Representation}
%{{{

% Johannes Ballé, Valero Laparra, and Eero P. Simoncelli. 2017. End-to-end Optimized Image Compression. In 5th International Conference on Learning Representations, ICLR 2017, Toulon, France, April 24-26, 2017

In ANN-based image compression, an autoencoder can be used to find the
latent representation of an (input) image. The compressor is the first
part of the autoencoder and the decompressor is the second
part. During compression, the latent representation is quantized
and the entropy compressed.

%}}}

\end{comment}
