% Emacs, this is -*-latex-*-

\input{../../definitions}
\title{\SM{} - Study Guide - Milestone 5 - Task 1 - Removing Spatial Redundancy with the 2D-Discrete Cosine Transform}

\maketitle

\tableofcontents

\section{Description of the task}

This task shows the effect in the RD
curve~\cite{vruiz__information_theory} of using the
\href{https://en.wikipedia.org/wiki/Discrete_cosine_transform}{DCT}~\cite{vruiz__DCT}
when it is applied by 2D blocks over the space domain of an image. The
2D-DCT is the key of the spatial decorrelation (information
concentration) used in some of the most relevant image and video
compressors, such as \href{https://en.wikipedia.org/wiki/JPEG}{JPEG}
and \href{https://en.wikipedia.org/wiki/Advanced_Video_Coding}{H.264},
\href{https://en.wikipedia.org/wiki/Advanced_Video_Coding}{HEVC} and
\href{https://en.wikipedia.org/wiki/Versatile_Video_Coding}{VVC}.

\subsection{Spatial redundancy and 2D-DCT decorrelation}

All images, except the noise\footnote{If the noise is also white, then
even no statistical redundancy should be present}, exhibit some degree
of spatial redudancy, that can be recognized because the color of
neighbour pixels tends to be similar or some kind of pattern is
repeated in the image. Such correlation can be exploited by the 2D-DCT to
concentrate information in a small number of (DCT)
coefficients~\cite{vruiz__DCT}.

\subsection{RD gain provided by the 2D-DCT}

The improvement resulting of the use of the 2D-DCT can be
quantified if we compare the RD curves considering or not the
transform.\footnote{This kind of comparison is commonly called
``\emph{Ablation Study}''.} The RD The visual information of an image can be
concentrated in a small number of DCT coefficients using the following
algorithm:
\begin{enumerate}
\item Divide the image into blocks. Usually, the blocks are not
  overlapping, but this is not mandatory because overlapped blocks
  could minimize the visually-perceptible blocking artifacts generated
  at low bit-rates. Notice also that the block size can be adapted to
  the local characteristics of the image with the aim of minimizing
  the RD curve (for example, in those regions where the RD performance
  of the 2D-DCT is small, the block size should be also small).
\item Transform each block using the 2D-DCT to each block-component,
  generating a subband-component~\cite{vruiz__transform_coding} per
  block-component.
\item Quantize and compress each $i$-th block with a quantization step
  size ${\mathbf\Delta}_i$ such that the overall RD is minimized. To
  achieve this, the selected RD slope used in each subband-component
  should be the same~\cite{vetterli2014foundations}. Notice that,
  because the DCT is orthonormal, the distortion (generated by the
  quantization) is not affected by the 2S-DCT and therefore, the
  distortion in the reconstructed image can be computed directly in
  the 2D-DCT domain.
\item Compress the quantization indexes of all the
  subband-components. For the sake of simplicity, we can consider the
  indexes as ``pixels'' and use a lossless image compressor, such as
  PNG. In addition, all the information necessary to reconstruct the image
  (such as the block partitioning pattern or the selected
  ${\mathbf\Delta}$) should be encoded in the code-stream.
\end{enumerate}

\section{What do I have to do?}
Run the \href{https://github.com/Sistemas-s-Multimedia.github.io/blob/master/contents/2D_DCT/RD_performance.ipynb}{notebook}.
%With the aim of checking that distortion can be measured in both, the
%transform and the image (RGB) domain, extend the
%previous \href{https://github.com/Sistemas-s-Multimedia.github.io/blob/master/milestones/07-%DCT/block_DCT_compression.ipynb}{notebook}
%to compute the distortion in the RGB domain. Check that the result is
%the same.

\section{Timming}

Please, finish this milestone before the next class session.

\section{Deliverables}

None.

\section{Resources}

\renewcommand{\addcontentsline}[3]{}% Remove functionality of \addcontentsline
\bibliography{maths,data-compression,signal-processing,DWT,image-compression,image-processing}
