% Emacs, this is -*-latex-*-

\input{../../definitions}
\title{\SM{} - Removing $\text{RGB}$ Redundancy with the DCT (Discrete Cosine Transform)}

\maketitle
\tableofcontents

\section{The RGB Domain}

\subsection{Objective}
The image compressor developed in the task
\href{https://sistemas-multimedia.github.io/contents/gray_SQ/}{Scalar
  Quantization of Gray-scale Images} can be extended to encode
$\text{RGB}$ (Red, Green and Blue) color images if each color channel
is quantized~\cite{vruiz__scalar_quantization} and the quantization
indexes are entropy encoded (for example, with
PNG~\cite{vruiz__PNG}). Notice that, as in the task
\href{https://sistemas-multimedia.github.io/contents/gray_SQ/}{Scalar
  Quantization of Gray-scale Images}, we are again exploiting the fact
that, in general, humans are not very proficient recognizing a large
number of simultaneous colors.\footnote{This can be classified as a
type of visual redundancy.}

\subsection{Scalar quantization applied to the $\text{RGB}$ domain}
The $\text{RGB}$ domain is additive~\cite{burger2016digital}, which
means that the distortion generated in one of the channels is
identical to the distortion (for example, in terms of the RMSE)
generated in the reconstructed image. Therefore, the set of
quantization patterns used in the compressor is
\begin{equation}
  \mathbf{\Delta}^{\text{R}}_i = \mathbf{\Delta}^{\text{G}}_i =
  \mathbf{\Delta}^{\text{B}}_i,
  \label{eq:simple_Q}
\end{equation}
where $i$ represents the $i$-th quantization step size.  Notice also
that any other combination of quantization steps should generate some
bias of the chroma towards one of the primary colors (red, green and
blue).

\subsection{Optimal quantization pattern}
\label{sec:optimal_RD}
The quantization pattern defined in the Eq.~\eqref{eq:simple_Q} will
be optimal only if for all the quantization steps, the condition
\begin{equation}
  \mathbf{\lambda}^{\text{R}}_i = \mathbf{\lambda}^{\text{G}}_i =
  \mathbf{\lambda}^{\text{B}}_i
  \label{eq:optimal_lambda}
\end{equation}
is satisfied. In the previous
\href{https://github.com/Sistemas-Multimedia/Sistemas-Multimedia.github.io/blob/master/contents/RGB_SQ/RGB_SQ.ipynb}{notebook}
we can explore (at least, visually) the grade of compilance of
Eq.~\eqref{eq:optimal_lambda}.

\subsection{Scalar quantization in the $\text{RGB}$ domain}

Supposing that we are using a dead-zone quantizer, a $\text{RGB}$
image can be quantized, channel by channel, using quantization steps
$\mathbf{\Delta}_{\text{R}}$, $\mathbf{\Delta}_{\text{G}}$, and
$\mathbf{\Delta}_{\text{B}}$ (being $\mathbf{\Delta}$ the
``quantization pattern''). A reasonable question that arises here is:
given a target bit-rate $R$ for the compressed frame, how the
quantization steps should be chosen to minimize the distortion?

At this point we can consider two different quantization
perspectives. In the first one, if we consider a strictly visual
effect, any alternative different from
\begin{equation}
  \mathbf{\Delta}_{\text{R}} = \mathbf{\Delta}_{\text{G}} = \mathbf{\Delta}_{\text{B}}
  \label{eq:simple_Q}
\end{equation}
will produce some alteration in the color (also called the
\href{https://en.wikipedia.org/wiki/Chrominance}{``chroma''}) of the
reconstructed image.

A second perspective considers only a pure
\href{https://en.wikipedia.org/wiki/Rate-distortion_theory}{Rate/Distortion
  (RD) performance}~\cite{vruiz__information_theory}. The $\text{RGB}$
color domain is additive, which means that the distortion generated in
the quantized image is the sum of the distortions generated in each of
the channels, when they are quantized independently. Therefore, the
optimal RD curve for the reconstructed image (using the $3$ channels)
can be obtained through the steps:
\begin{enumerate}
\item Compute the RD curve for each channel. Notice that here we must
  define a set of quantization steps for each channel. For simplicity,
  use the same set for the three channels.
\item Put all the RD points in a list and sort the list by the slopes
  of the points. Since the points have been ordered by their slope,
  the sequence of quantization patterns should generate the optimal
  curve.
\item Use the corresponding quantization patterns (defined by the points
  of the list) to compute the RD curve of the reconstructed image.
\end{enumerate}

This algorithm is independent of how the set of quantization steps
used in each channel has been selected. If the quantization steps used
to describe the RD curves of each channel are different, the points of
the definitive curve will be different, but not its shape or its
localization in the RD space.

If we are used the same quantization steps for each channel, the
points generated when
\begin{equation}
  \mathbf{\Delta}_{\text{R}} = \mathbf{\Delta}_{\text{G}} = \mathbf{\Delta}_{\text{B}}
\end{equation}
must occur in the definitive RD curve. Notice that the rest of points
of the curve should appear between these points, but do not
necessarily belong to the convex hull of the curve (altough they could
be used to perform a finer rate-control).

\subsection{Increasing the number of RD points (TO-DO)}
\label{sec:increasing}
If the Eq.~\eqref{eq:simple_Q} does not need to be satisfied, we can
increase the resolution of the rate/distortion control using also the
``intermediate'' points of the bit-plane progressive RD
curve. Therefore, starting for example by the (usually top) left point
of the RD curve, after each Eq.~\eqref{eq:simple_Q}-point, we must
decide which option is more advantageous: (1) encode the next
bit-plane of the $\text{R}$ channel, (2) encode the next bit-plane of
the $\text{G}$ channel, or (3) encode the next bit-plane of the
$\text{B}$ channel. We should select the option that provides the
highest RD slope~\cite{vruiz__information_theory}.


%%%%%%

\section{The Color DCT Domain}

\subsection{Objective}
This task shows how to exploit the inter-channel
redundancy~\cite{vruiz__visual_redundancy,vruiz__information_theory}
in $\text{RGB}$ images with the DCT~\cite{vruiz__DCT} applied over the
color dimension~\cite{vruiz__image_IO}. The DCT is one of the most
used transforms in audio, image and video coding, because it is
orthonormal (decorrelates, preserves the energy and all the DCT
subbands has $1$-gain)~\cite{vruiz__transform_coding} and, in general,
performs (decorrelates) quite well.

\subsection{Decorrelation with the DCT}
The DCT convolves cosine functions with the input sinal (in this case,
the components of the pixels of the input image). The output of the
transform is a set of three subbands $\{\text{DCT}^0, \text{DCT}^1,
\text{DCT}^2\}$. The number of coefficients in a subband
$\text{DCT}^s$ is the same that the number of $\text{RGB}$-pixels in
the original image $\mathbf{I}$, or the number of monochrome-pixels in
a channel $\mathbf{I}^c$ (see this
\href{https://github.com/Sistemas-Multimedia/Sistemas-Multimedia.github.io/blob/master/contents/RGB_DCT_SQ/RGB_DCT_SQ.ipynb}{notebook}).

As a result of the use of the DCT, in most of the images the energy is
concentrated in the low frequency
subband~\cite{vruiz__transform_coding} ($\text{DCT}^0$), which
basically represents the
\href{https://en.wikipedia.org/wiki/Luminance}{luminance} of the
image (see the \href{https://github.com/Sistemas-Multimedia/Sistemas-Multimedia.github.io/blob/master/contents/RGB_DCT_SQ/RGB_DCT_SQ.ipynb}{notebook}).

\subsection{Effects of the transform}
The concentration of the energy in the subband $\text{DCT}^0$
increases the dynamic range of the DCT-color domain compared to the
dynamic range of the original $\text{RGB}$ image. This has two
interesting consequences:
\begin{enumerate}
\item The entropy of the progressive encoding (by bit-planes) of the
  integer representation of the DCT coefficients is reduced (compared
  to the entropy of the bit-planes of the original $\text{RGB}$
  pixels). Therefore, the RD performance is improved because in terms
  of RMSE/bit-rate we are closer to the origin of coordinates.
\item The number of different quantization steps can be higher, and
  therefore it is possible to increases the number of points in the RD
  curve. This potentially increases the granularity of the RD control,
  when applied.
\end{enumerate}

\subsection{Scalar quantization applied to the DCT domain}
Since the DCT is orthonormal (orthogonal and unitary), we can use the
simple quantization pattern
\begin{equation}
  \mathbf{\Delta}^{\text{DCT}^0} = \mathbf{\Delta}^{\text{DCT}^1} = \mathbf{\Delta}^{\text{DCT}^2}.
  \label{eq:simple_Q}
\end{equation}
Therefore, there is not differente between quantizing in the
$\text{RGB}$ or the ($\text{DCT}^0,\text{DCT}^1,\text{DCT}^2$) color
domain, except that in the last case, we must consider one extra
bit-plane to encode.

\subsection{A better rate control (TO-DO)}
\label{sec:increasing}
As we did in the case of the
\href{https://sistemas-multimedia.github.io/contents/RGB_SQ/}{Scalar
  Quantization in the $\text{RGB}$ Color Domain}, we can also use the
intermediate RD points between each Eq.~\eqref{eq:simple_Q}-point to
increase the resolution of the decompressed image.

%%%%%%%%%%%%%%%%%

\section{Scalar versus vector quantization of RGB images}
%{{{

SQ (Scalar
Quantization)~\cite{vruiz__scalar_quantization,sayood2017introduction}
would be an optimal solution only if the image colors are uniformly
distributed within
\href{https://en.wikipedia.org/wiki/RGB_color_model}{the RGB
  cube}. However, the typical color distribution in natural images is
anything but uniform, with some regions of the color space being
densely populated and many potentially used colors entirely
missing. In this case, depending on the quantization step
size~\cite{vruiz__signal_quantization}, SQ is not optimal because the
used colors may not be sampled with suﬃcient density while at the same
time the encoding system is considering colors that do not appear in
the image at all~\cite{burger2016digital}.

On the other hand, VQ (Vector
Quantization)~\cite{vruiz__vector_quantization,sayood2017introduction}
applied to the color domain does not treat the individual RGB
components (also refered by
\href{https://en.wikipedia.org/wiki/Color_image}{channel}s) separately
as does scalar quantization, but each used color vector ${\mathbf C}_i
= ({\mathbf R}_i, {\mathbf G}_i, {\mathbf B}_i )$ in the image is
treated as a minimum structure. Starting from a set of original color
tuples ${\mathbf C} = \{{\mathbf C}_1, {\mathbf C}_2, \ldots ,{\mathbf
  C}_m\}$, where $m$ is the number of different colors found in the
image, the task of a vector quantizer in this context is to:
\begin{enumerate}
\item Find a set of $n$ representative color vectors (the so called
  \emph{code-book}) ${\mathbf C}' = \{{\mathbf C}'_1, {\mathbf C}'_2
  ,\ldots , {\mathbf C}'_n \}$, where $n < m$.
\item Replace each original color ${\mathbf C}_i$ by one of the new
  color vectors ${\mathbf C}'_j\in {\mathbf C}'$, where the resulting
  deviation from the original image shall be minimal.
\end{enumerate}

%}}}


\section{Resources}

\renewcommand{\addcontentsline}[3]{}% Remove functionality of \addcontentsline
\bibliography{maths,data-compression,signal-processing,DWT,image-compression,image-processing,information_theory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{comment}

Implement a lossy image compressor that exploits the inter-channel
redundancy~\cite{vruiz__information_theory} in $\text{RGB}$ images. To
achive this, we use the DCT~\cite{vruiz__DCT} applied over the color
dimension~\cite{vruiz__image_IO}, resulting in an concentration of the
energy of the image in the first
subband~\cite{vruiz__transform_coding} ($\text{DCT}^0$) which
basically represents the
\href{https://en.wikipedia.org/wiki/Luminance}{luminance} of the
image. Notice that the visual
redundancy~\cite{vruiz__visual_redundancy} is not considered.



\section{Distortion control}
We do not provide any accurate distortion control algorithm.  After
using the DCT, the coefficients are quantized with a deadzone
quantizer, simulating a bit-plane encoding. However, the dynamic range
of the $\text{DCT}^0$ coefficients doubles the dynamic range of the
pixels, and therefore, the highest quantization step will be
$\Delta=256$. Therefore, we will have eight points in our RD curve
that satisfy that
\begin{equation}
  \mathbf{\Delta}_{\text{DCT}^0} = \mathbf{\Delta}_{\text{DCT}^1} = \mathbf{\Delta}_{\text{DCT}^2}.
  \label{eq:simple_Q}
\end{equation}

Notice that the DCT is orthogonal and therefore, the contributions of
the subbands (to the quality of the reconstructed image) are
independent and therefore, additive. Because the DCT is also
orthonormal, we can measure the distortion in both, the image and the
transform domain.

\section{Rate control}
Since PNG does not provide accurate rate control, we can only select
between eight bit-rates that safisfy the Eq.~\eqred{eq:simple_Q}.

\section{Scalar quantization in the ($\text{DCT}_0,\text{DCT}_1,\text{DCT}_2$) domain}
%{{{

The DCT is orthonormal, and thereore, the distortion can be measured
directly in the transform domain. This means that we can treat the DCT
coefficients as if would be pixels. Therefore, an optimal algorithm
for RDO~\cite{vruiz__information_theory} is:

\begin{enumerate}
\item Find the RD curve for each subband. We will use the same
  quantization steps for all the subbands. This action generates three
  curves.
\item Join all the RD points in a single list and sort it by the slope
  of the
  points~\cite{vruiz__information_theory}. Track\footnote{Storing with
  the point the combination subband index and quantization step used
  for each point.} the quantization steps used for each point.
\item Use the quantization patters described by the points of the
  sorted list to obtain the definitive RD curve. Notice that, since
  the DCT is orthonormal, we could estimate accurately the distortion
  in the definitive curve using directly the distortions of the
  subbands.
\end{enumerate}

Notice that the points that satisfy that
\begin{equation}
  \mathbf{\Delta}_0 = \mathbf{\Delta}_1 = \mathbf{\Delta}_2
  \label{eq:deltas}
\end{equation}
should belong to the definitive RD curve and should describe also the
convex hull of the curve. There is not guarantee that the rest of
points in which the Eq.~\eqref{eq:deltas} is not true will belong to
the convex hull (although they could be used to perform a finer
rate-control).

This algorithm has been implemented in this \href{https://github.com/Sistemas-Multimedia/Sistemas-Multimedia.github.io/blob/master/contents/RGB_DCT/RGB_DCT.ipynb}{notebook}.

\section{Ablation study}
The improvement resulting from the use of the DCT applied over the RGB
domain can be quantified if we compare the RD curves considering or
not its use.\footnote{This kind of comparison is commonly called
``\emph{Ablation Study}'' because we are ''removing'' some
functionality of the system to see how its ausence impacts on the
performance.} 

\end{comment}

\begin{comment}
   The steps
of this procedure are (see the
\href{https://github.com/Sistemas-Multimedia/Sistemas-Multimedia.github.io/blob/master/contents/color_DCT/RGB_DCT.ipynb}{notebook}):
\begin{enumerate}
\item  This fact allows us to use a ``fast'' (with
  linear
  \href{https://en.wikipedia.org/wiki/Computational_complexity}{complexity})
  RDO (Rate/Distortion Optimization) algorithm to find which
  combination of quantization steps $\mathbf{\Delta}$ that
  ``minimizes'' the RD curve of the reconstructed image, which can be
  computed as the sum\footnote{Remember, the constribution of the
  subbands are independent and additive.} of the RD curves generated
  by the $3$ subbands ($\text{DCT}^0$, $\text{DCT}^1$, and
  $\text{DCT}^2$). Therefore, we can:
  \begin{enumerate}
  \item Quantize and compress\footnote{We can consider that the
  coefficients are gray-scale pixels and use a image compressor.} each
    subband with a quantization step $\mathbf{\Delta}_s$. This results
    in $3$ RD curves, one per subband. We define the slope of the
    $n$-th point in a RD curve as the slope of the straight line
    passing through the points $n$ and $n+1$ (see the
    Fig.~\ref{fig:slope_computation}). The slope of the point for the
    highest rate is defined to $0$.  The results generated by this
    step are $3$ RD curves (one per subband).
  \item Put all the RD points in a list, tracking\footnote{Remembering
  the combination subband index and quantization step used for each
  point.} the quantization pattern used for each RD point. Compute the
    list of points of the RD curve of the reconstructed image as the
    (in descending order) sorted list, by slope, of RD points.
  \item (Optional) If the image compressor used in the first step
    generates less data overhead when a color image is compressed
    considering the three channels at the same time, recompute the
    rate of each RD point of the definitive curve. For example, in the
    case of PNG~\cite{vruiz__PNG}, even when the intercomponent
    redundancy is not exploited, the number of headers is decreased if
    we perform only one compression per image.
  \end{enumerate}

  The orthonormality of the DCT also allows us to compute the
  definitive distortion directly in the DCT domain.

  Notice also that this algorithm has been designed considering the
  fact that, ideally (where the RD curve is continuous), the rate
  selected for each subband should be the
  same~\cite{vruiz__information_theory,vetterli2014foundations}, and
  therefore, should generate a RD curve such that the sum of the
  distances of the points of the RD curve to the point $(0,0)$ is
  minimized.
  
\end{enumerate}

As a final remark, take into consideration that the quantization steps
used in each subband should be selected considering aspects such as
the ganularity of the rate-control\footnote{How many points has the RD
curve.} and the features of the decoding process\footnote{For example, if we
want to provide progressive bit-plane decoding, the quantization steps
should be powers of $2$.}

\section{An even faster (but coarser) rate-control algorithm}

The running time of the previous algorithm depends on the number of
subbands, that in our case is only $3$. However, we can develop a
faster rate-control procedure (avoiding to compute the RD curves of
the subbands that is the heaviest part) if we suppose that the RD
slopes are basically determined by the distortion (i.e., without
considering the rate)\footnote{Notice that this basically means that
all the subband's curves have the
same \footnote{https://en.wikipedia.org/wiki/Domain_of_a_function}{domain}
and that the RD points occurs at the same rates}.

Obviously, in the practice it is very unlikely to happen, but if the
shapes of the curves are close enough, we can suppose that the slopes
are the same for the same points in the different curves. In the case
of the DCT that is orthonormal (the gains of the forward and the
backward filters is $1$), this is the same that define the
quantization pattern
\begin{equation}
  \mathbf{\Delta}_0 = \mathbf{\Delta}_1 = \mathbf{\Delta}_2.
\end{equation}

Notice that one way to check if this quantization algorithm is optimal
(for the obtained RD points) is to check if the ``slower'' algorithm
outputs such quantization patterns. If this is true, the ``faster''
rate-control algorithm is optimal, at least for the encoded image.

\end{comment}
